Predicting Fire Risk: \\ An Evaluation of the Fire Danger Index

### 2016-03-09 paper.tex
Mostly formatting

% !TeX spellcheck = <none>
\documentclass[11pt,a4paper]{article}

\usepackage{fullpage,booktabs,amsmath,paralist,parskip}
\usepackage[section]{placeins}
\usepackage{gensymb,microtype,todonotes}
\usepackage{natbib}
\usepackage{pdfpages}
\usepackage{listings}

@@

%% GRAPHICS
\usepackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

@@

\subsection{Predicting fire activity}

@@

The CFA divides the state into different districts for management purposes. I use these districts to capture different terrain types, fuel types and population density in Victoria. Dividing the data into these districts allows us to observe any difference in the behaviour of the variables across different parts of the state. Figure~\ref{fig:dist} shows a map of the districts and their labels.

\begin{figure}
  \centering
  \includegraphics[width=1\columnwidth]{figures/cfadis.pdf}
  \caption{Country Fire Authority Districts}
  \label{fig:dist}
\end{figure}

@@

In this thesis, I predict probability of ignition using McArthur's Forest Fire Danger Index (FFDI), the meteorological variables used as inputs for the FFDI, grassland curing, vapour pressure at 9am and rainfall.

@@

%Graphs, need to redo better
\begin{figure}
  \centering
  	\includegraphics[width=1\columnwidth]{figures/silvan.pdf}
  \caption{Time series of meteorological variables for Silvan (-37.82, 145.42)}
  \label{fig:silv}
\end{figure}

@@

%WS, Temp and FDR contours, see if can change colours to match with the diagram
\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/tw_fdr.pdf}
  \caption{Wind Speed and temperature coloured by fire danger rating}
  \label{fig:tw_fdr}
\end{figure}

@@

The grassland curing data were obtained through the Bureau of Meteorology TERN server and AusCover (http://www.auscover.org.au). AusCover is the remote sensing data products facility of the Terrestrial Ecosystem Research network (TERN, http://www.tern.org.au). There are several different satellite models available to provide curing values for image pixels across the state of Victoria. I use `MapVictoria' \citep{martin15} as this is the current algorithm used by the CFA and is shown to outperform the other curing models.

@@

\begin{figure}[p]
\includegraphics[width=1\columnwidth]{figures/curdist.pdf}
  \caption{Grassland curing evolution over time }
  \label{fig:cur_evol}
\end{figure}

% level plot of GS curing, probably needs resizing
\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/cur_lp.pdf}
  \caption{Level Plot of Grassland Curing }
  \label{fig:cur_lp}
\end{figure}

@@

Vapour pressure, in meteorology, describes the partial pressure of water vapour in the atmosphere, it is directly related to the number of water vapour molecules in the air.  \citet{harris14} find there is a significant relationship between fire activity and vapour pressure among other climate variables. I investigate whether vapour pressure at 9am can improve prediction of fire occurrence.

@@

% level plot of vapour pressure, level plot may be unnesc could just have std plot
\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/vap_ras.pdf}
  \caption{Raster plot of vapour pressure 9am }
  \label{fig:vap_ras}
\end{figure}

@@

%Commented out until final printing,
%Map of fire ignitions in Victoria
\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/firemap.pdf}
  \caption{Map of fire ignition 2000 - 2009}
  \label{fig:firemap}
\end{figure}

%timeseries histogram with count of ignitions per fortnight
\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/occurence.pdf}
  \caption{Number of ignitions by fortnight}
  \label{fig:occ}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/tw_fire.pdf}
  \caption{Wind Speed and temperature coloured by fire/no fire}
  \label{fig:tw_fire}
\end{figure}

Figure~\ref{fig:tw_fire} is constructed in the same manner as Figure~\ref{fig:tw_fdr} but instead of grouping by the fire danger rating, I group by Fire/No Fire. Windspeed and temperature play the largest role in determining the FFDI yet there is no discernible difference between the distributions of the two variables when there is a fire compared to no fires. This is the first indicator that the FFDI alone may not be a good predictor of probability of ignition.

\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/dh_fire.pdf}
  \caption{Drought factor and relative humidity coloured by fire/no fire}
  \label{fig:dh_fire}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/gv_fire.pdf}
  \caption{Grassland curing and vapour pressure coloured by fire/no fire}
  \label{fig:gv_fire}
\end{figure}

@@

\subsection{Assembling datasets}

@@

The first step was to reduce the dataset to a more manageable size. I converted the NCDF file to a data table using the raster and data.table packages in R \citep{raster, datatable}. I used ESRI shapefiles of Australian Geographical Classification boundaries \citep{ABS1259} to sort the grid points into states and selected only grid points in Victoria. The datasets intersect between 2000 and 2009, subsetting the WFR data reduces the size of the dataset to just over 8 million data points.

From Figure~\ref{fig:occ} it is clear the majority of bushfires occur during the fire season. I reduce the size of the dataset, taking only observations in Victorian fire season (October to March) for a number of reasons. First, from a policy perspective the FFDI, fire danger rating systems, grassland curing are only calculated during the fire season so using these variables outside the time periods they are available does not make intuitive sense. From a practical perspective, this paper is looking at predicting the types bushfires that threaten lives and property which we know to occur in the summer months.

@@

From the fire occurrence dataset this study used the date and coordinates of the fire. The coordinates were given in the Australian Map Grid (AMG66), a coordinate reference system derived from a Universal Transverse Mercator(UMT) projection of latitudes and longitudes on the Australian Geodetic Datum (AGD) \citep{featherstone96}. To merge the meteorological data set and the fire occurrence dataset I needed to convert from this UMT projection to longitude and latitude using the sp package in R \citep{sp08}.

Once the coordinates of fires were converted to the same projection I matched each fire coordinate to its nearest neighbour from the meteorological dataset. Once matched it was straightforward to merge the two datasets based on grid point and date \citep{datatable}.

The different CFA districts and regions were added into the dataset using the raster package and another ESRI shapefile containing the district boundaries \citep{raster}.
Digital boundaries for the CFA area compared to MFB or DELWP response areas were unavailable, as such an alternate method to subset geographical area was used. I made the assumption that if the CFA had not attended a bushfire between 2000 and 2014 within a 5km region of the grid point, then the grid point was unlikely to be in a CFA area. Accordingly any grids that had no fire occur in them since 2000 were excluded from the dataset.

The three additional variables (grassland curing, vapour pressure and precipitation) came as raster files each of higher resolutions than the original 10km grid. In all three sets we have some missing values or faulty readings due to technological fault. To minimise the impact of these for each for each grid point in the meteorological dataset I find the pixels contained in each grid point and take the median value for each time period. The functionality to combine spatial points and raster layers is provided by the raster package \citep{raster}. For any remaining missing values I make the assumption that cloud cover or faulty barometer readings are sufficiently random such that missing values will not bias results.

@@

\subsection{Logistic regression}

Use of logistic models is the usual method for dealing with a binary response problem and it is common in studies on fire occurrence or ignition.  In this project I use logistic regression models and in addition to the linear constructions used in \citet{andrews03} and \citet{del11} I take into account the non-linear relationship believed to exist between the explanatory variables and dependent variable through the inclusion of regression splines.

@@

\subsection{Classification trees}

@@

To approach over fitting, one of two different approaches can be taken: either a rule can be applied that prevents the tree growing too large (for example by setting a threshold on the minimum amount of information gain required to produce a split), or the tree can be allowed to grow large and pruning can be applied afterwards to produce a subtree. In this paper I use 10-fold cross validation on the training data to choose the precision parameter for the classification tree.

Decision trees tend to be very sensitive to the sample used to grow them. Taking this into account I build a number of trees using different balanced samples of the data and compare the output. An alternative solution is random forests (discussed below) as they have the additional advantage of reducing the problem of high variance which plagues decision trees \citep{james13}.

@@

\subsection{Random forests}

A random forest is a technique that uses classification trees as a key building block and helps address the limitations of classification trees and improve predictive performance. They involve multiple trees and random samples of our predictors \citep{berk08}.
The general algorithm is as follows for our dataset of N observations and P predictor variables \citep{berk08, varian14}:

\begin{enumerate}
  \item Take a bootstrap sample (random sample of size N with replacement) from the data.
  \item Take a random sample of size $M < P$ (typically $M \approx \sqrt{P}$, \cite{james13}, without replacement) from the predictors.
  \item Construct a classification tree using these samples of data and predictors.
  \item Repeat from step 2 for each subsequent split until the tree is as large as desired. Do not prune.
  \item Drop observations not included in the bootstrap sample from step 1 (known as out-of-bag data) down the tree. The class assigned to each observation is stored along with each observations predictor values.
  \item Repeat steps 1-5 a large number of times to grow a forest of trees.
  \item In order to determine the classification of a new observation. Have each tree make a classification and use a majority vote for the final prediction.
\end{enumerate}

@@

I expect the random forest model to give improved out-of-sample fit particularly with highly non-linear data and despite being difficult to interpret, the random forests can offer insight regarding which predictors contribute the biggest improvements to prediction accuracy \citep{varian14}.

\subsection{Classification with unbalanced classes}

@@

I address the class imbalance problem in two ways, first in building the model I create balanced samples by under sampling the `no fire' class in a similar manner to \citep{padilla11}.  The second approach is to assess each model's performance in predicting the test data using `balanced accuracy' rather than overall accuracy \citep{mosley13, japkowicz00}.

Under-sampling the no fire class addresses the problem of imbalanced classes but I also use it to reduce the serial and spatial correlation that is present in the dataset by using a `near-maximin' sampling method thereby ensuring all the data points in the sample are separate from each from each other in space and time to reduce the impact of correlation.

For consistency all three of the classification techniques are estimated on the balanced training set and evaluated by  on the test data. I evaluate based on receiver operating characteristic (ROC) plots  and using $2\times2$ confusion matrices.

@@

As a secondary measure of model quality, I use ROC plots as they remove the requirement of assuming a cut-off value.

An ROC plot shows the change in true positive rate (sensitivity) and false positive rate (1- specificity) for different probability thresholds. For each model I plot the ROC and calculate the area under the curve (AUC) as a measure of the model performance. The AUC measures a classification model's performance, ranging from 0.5 (random) to 1 (perfect accuracy).

@@

The baseline model is given by a simple logistic regression of fire on FFDI, I then add the inputs to the FFDI and  the other variables (curing, vapour pressure and rainfall).

@@

%Do I need to show output for this?

@@

I add splines to Model 3 in stages using the confusion matrices for 10-fold cross validation to assess accuracy \citep[][using the caret package by]{caret}. The best performing model based on balanced accuracy and number of false negatives (incorrectly predicted `No Fires') is reported in Table~\ref{table:logf}.

@@

%ROC for logits
\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/all_roc.pdf}
  \caption{ROC curve for logistic models}
  \label{fig:ffdi}
\end{figure}

@@

As classification trees can be sensitive to the sample used, I constructed several trees on different balanced samples of the data. As each of the trees were structured very similarly I only report two in Figure~\ref{fig:tree2}.

% Trees with all variables
\begin{figure}
  \centering
	\includegraphics[page=1, width=1\columnwidth]{figures/trees2.pdf}
	\includegraphics[page=2, width=1\columnwidth]{figures/trees2.pdf}
  \caption{Trees with all variables}
  \label{fig:tree2}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/trees_roc.pdf}
  \caption{ROC curve for Tree Models}
  \label{fig:troc}
\end{figure}

\subsection{Random forest}

I produce the random forest using the caret and randomForest packages \citep{caret, rf} in R. The random forest is built using the balanced training set and then assessed on its prediction on the test set.

@@

\begin{figure}
  \centering
  	\includegraphics[width=1\columnwidth]{figures/varimp.pdf}
  \caption{Variable Importance for random forest}
  \label{fig:vimp}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=1\columnwidth]{figures/roc_rf.pdf}
  \caption{Variable Importance for random forest}
  \label{fig:rocrf}
\end{figure}

@@

This thesis takes a very simplistic view of the dataset, rather than utilising the spatial and time series component, I treat each spatial and time point as independent. Utilising the lagged variables will no doubt improve the model for example we'd expect after a week of hot windy days the probability of ignition would be higher compared to a week of cold days or rain.

@@

From the different models in this paper, I conclude that the FFDI alone of a poor predictor of bushfire ignition.

>>>>

% !TeX spellcheck = <none>
\documentclass[11pt,a4paper]{article}
\usepackage[a4paper,text={17cm,25cm}]{geometry}

\usepackage{booktabs,amsmath,paralist,parskip,hyperref}
\usepackage[section]{placeins}
\usepackage{gensymb,microtype,todonotes}
\usepackage{natbib}
\usepackage{pdfpages}
\usepackage{listings}

@@

%% GRAPHICS
\usepackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}
\usepackage[section]{placeins}

@@

\subsection{Predicting Fire Activity}

@@

The CFA divides the state into different districts for management purposes. We use these districts to capture different terrain types, fuel types and population density in Victoria. Dividing the data into these districts allows us to observe any difference in the behaviour of the variables across different parts of the state. Figure~\ref{fig:dist} shows a map of the districts and their labels.

\begin{figure}
  \centering
  \includegraphics[width=.7\textwidth]{figures/cfadis.pdf}
  \caption{Country Fire Authority Districts}
  \label{fig:dist}
\end{figure}

@@

We predict probability of ignition using McArthur's Forest Fire Danger Index (FFDI), the meteorological variables used as inputs for the FFDI, grassland curing, vapour pressure at 9am and rainfall.

@@

%Graphs, need to redo better
\begin{figure}
  \centering
  	\includegraphics[width=\textwidth]{figures/silvan.pdf}
  \caption{Time series of meteorological variables for Silvan (-37.82, 145.42)}
  \label{fig:silv}
\end{figure}

@@

%WS, Temp and FDR contours, see if can change colours to match with the diagram
\begin{figure}
  \centering
	\includegraphics[width=\textwidth]{figures/tw_fdr.pdf}
  \caption{Wind Speed and temperature coloured by fire danger rating}
  \label{fig:tw_fdr}
\end{figure}

@@

The grassland curing data were obtained through the Bureau of Meteorology TERN server and AusCover (http://www.auscover.org.au). AusCover is the remote sensing data products facility of the Terrestrial Ecosystem Research network (TERN, http://www.tern.org.au). There are several different satellite models available to provide curing values for image pixels across the state of Victoria. We use `MapVictoria' \citep{martin15} as this is the current algorithm used by the CFA and is shown to outperform the other curing models.

@@

\begin{figure}[p]\centering
\includegraphics[width=.95\textwidth]{figures/curdist.pdf}
  \caption{Grassland curing evolution over time }
  \label{fig:cur_evol}
\end{figure}

% level plot of GS curing, probably needs resizing
\begin{figure}
  \centering
	\includegraphics[width=.8\textwidth]{figures/cur_lp.pdf}
  \caption{Level Plot of Grassland Curing }
  \label{fig:cur_lp}
\end{figure}

@@

Vapour pressure, in meteorology, describes the partial pressure of water vapour in the atmosphere, it is directly related to the number of water vapour molecules in the air.  \citet{harris14} find there is a significant relationship between fire activity and vapour pressure among other climate variables. We investigate whether vapour pressure at 9am can improve prediction of fire occurrence.

@@

% level plot of vapour pressure, level plot may be unnesc could just have std plot
\begin{figure}
  \centering
	\includegraphics[width=.8\textwidth]{figures/vap_ras.pdf}
  \caption{Raster plot of vapour pressure 9am }
  \label{fig:vap_ras}
\end{figure}

@@

%Commented out until final printing,
%Map of fire ignitions in Victoria
\begin{figure}
  \centering
	\includegraphics[width=.8\textwidth]{figures/firemap.pdf}
  \caption{Map of fire ignition 2000 - 2009}
  \label{fig:firemap}
\end{figure}

%timeseries histogram with count of ignitions per fortnight
\begin{figure}
  \centering
	\includegraphics[width=.8\textwidth]{figures/occurence.pdf}
  \caption{Number of ignitions by fortnight}
  \label{fig:occ}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=\textwidth]{figures/tw_fire.pdf}
  \caption{Wind Speed and temperature coloured by fire/no fire}
  \label{fig:tw_fire}
\end{figure}

Figure~\ref{fig:tw_fire} is constructed in the same manner as Figure~\ref{fig:tw_fdr} but instead of grouping by the fire danger rating, we group by Fire/No Fire. Windspeed and temperature play the largest role in determining the FFDI yet there is no discernible difference between the distributions of the two variables when there is a fire compared to no fires. This is the first indicator that the FFDI alone may not be a good predictor of probability of ignition.

\begin{figure}
  \centering
	\includegraphics[width=\textwidth]{figures/dh_fire.pdf}
  \caption{Drought factor and relative humidity coloured by fire/no fire}
  \label{fig:dh_fire}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=\textwidth]{figures/gv_fire.pdf}
  \caption{Grassland curing and vapour pressure coloured by fire/no fire}
  \label{fig:gv_fire}
\end{figure}

@@

\subsection{Assembling Datasets}

@@

The first step was to reduce the dataset to a more manageable size. We converted the NCDF file to a data table using the raster and data.table packages in R \citep{raster, datatable}. We used ESRI shapefiles of Australian Geographical Classification boundaries \citep{ABS1259} to sort the grid points into states and selected only grid points in Victoria. The datasets intersect between 2000 and 2009, subsetting the WFR data reduces the size of the dataset to just over 8 million data points.

From Figure~\ref{fig:occ} it is clear the majority of bushfires occur during the fire season. We reduce the size of the dataset, taking only observations in Victorian fire season (October to March) for a number of reasons. First, from a policy perspective the FFDI, fire danger rating systems, grassland curing are only calculated during the fire season so using these variables outside the time periods they are available does not make intuitive sense. From a practical perspective, this paper is looking at predicting the types bushfires that threaten lives and property which we know to occur in the summer months.

@@

From the fire occurrence dataset this study used the date and coordinates of the fire. The coordinates were given in the Australian Map Grid (AMG66), a coordinate reference system derived from a Universal Transverse Mercator(UMT) projection of latitudes and longitudes on the Australian Geodetic Datum (AGD) \citep{featherstone96}. To merge the meteorological data set and the fire occurrence dataset we needed to convert from this UMT projection to longitude and latitude using the sp package in R \citep{sp08}.

Once the coordinates of fires were converted to the same projection we matched each fire coordinate to its nearest neighbour from the meteorological dataset. Once matched it was straightforward to merge the two datasets based on grid point and date \citep{datatable}.

The different CFA districts and regions were added into the dataset using the raster package and another ESRI shapefile containing the district boundaries \citep{raster}.
Digital boundaries for the CFA area compared to MFB or DELWP response areas were unavailable, as such an alternate method to subset geographical area was used. We made the assumption that if the CFA had not attended a bushfire between 2000 and 2014 within a 5km region of the grid point, then the grid point was unlikely to be in a CFA area. Accordingly any grids that had no fire occur in them since 2000 were excluded from the dataset.

The three additional variables (grassland curing, vapour pressure and precipitation) came as raster files each of higher resolutions than the original 10km grid. In all three sets we have some missing values or faulty readings due to technological fault. To minimise the impact of these for each for each grid point in the meteorological dataset we find the pixels contained in each grid point and take the median value for each time period. The functionality to combine spatial points and raster layers is provided by the raster package \citep{raster}. For any remaining missing values we make the assumption that cloud cover or faulty barometer readings are sufficiently random such that missing values will not bias results.

@@

Use of logistic models is the usual method for dealing with a binary response problem and it is common in studies on fire occurrence or ignition.  In this project we use logistic regression models and in addition to the linear constructions used in \citet{andrews03} and \citet{del11} we take into account the non-linear relationship believed to exist between the explanatory variables and dependent variable through the inclusion of regression splines.

@@

@@

To approach over fitting, one of two different approaches can be taken: either a rule can be applied that prevents the tree growing too large (for example by setting a threshold on the minimum amount of information gain required to produce a split), or the tree can be allowed to grow large and pruning can be applied afterwards to produce a subtree. In this paper we use 10-fold cross validation on the training data to choose the precision parameter for the classification tree.

Decision trees tend to be very sensitive to the sample used to grow them. Taking this into account we build a number of trees using different balanced samples of the data and compare the output. An alternative solution is random forests (discussed below) as they have the additional advantage of reducing the problem of high variance which plagues decision trees \citep{james13}.

@@

\subsection{Random Forests}

A random forest is a technique that uses classification trees as a key building block and helps address the limitations of classification trees and improve predictive performance. They involve multiple trees and random samples of our predictors \citep{berk08}.
The general algorithm is as follows for our dataset of N observations and P predictor variables \citep{berk08, varian14}:
\begin{enumerate}
  \item Take a bootstrap sample (random sample of size N with replacement) from the data.
  \item Take a random sample of size $M < P$ (typically $M \approx \sqrt{P}$, \cite{james13}, without replacement) from the predictors.
  \item Construct a classification tree using these samples of data and predictors.
  \item Repeat from step 2 for each subsequent split until the tree is as large as desired. Do not prune.
  \item Drop observations not included in the bootstrap sample from step 1 (known as out-of-bag data) down the tree. The class assigned to each observation is stored along with each observations predictor values.
  \item Repeat steps 1-5 a large number of times to grow a forest of trees.
  \item In order to determine the classification of a new observation. Have each tree make a classification and use a majority vote for the final prediction.
\end{enumerate}

@@

We expect the random forest model to give improved out-of-sample fit particularly with highly non-linear data and despite being difficult to interpret, the random forests can offer insight regarding which predictors contribute the biggest improvements to prediction accuracy \citep{varian14}.

\subsection{Classification with Unbalanced Classes}

@@

We address the class imbalance problem in two ways, first in building the model we create balanced samples by under sampling the `no fire' class in a similar manner to \citep{padilla11}.  The second approach is to assess each model's performance in predicting the test data using `balanced accuracy' rather than overall accuracy \citep{mosley13, japkowicz00}.

Under-sampling the no fire class addresses the problem of imbalanced classes but we also use it to reduce the serial and spatial correlation that is present in the dataset by using a `near-maximin' sampling method thereby ensuring all the data points in the sample are separate from each from each other in space and time to reduce the impact of correlation.

For consistency all three of the classification techniques are estimated on the balanced training set and evaluated by  on the test data. We evaluate based on receiver operating characteristic (ROC) plots  and using $2\times2$ confusion matrices.

@@

As a secondary measure of model quality, we use ROC plots as they remove the requirement of assuming a cut-off value.

An ROC plot shows the change in true positive rate (sensitivity) and false positive rate (1- specificity) for different probability thresholds. For each model we plot the ROC and calculate the area under the curve (AUC) as a measure of the model performance. The AUC measures a classification model's performance, ranging from 0.5 (random) to 1 (perfect accuracy).

@@

The baseline model is given by a simple logistic regression of fire on FFDI, we then add the inputs to the FFDI and  the other variables (curing, vapour pressure and rainfall).

@@

%Do we need to show output for this?

@@

We add splines to Model 3 in stages using the confusion matrices for 10-fold cross validation to assess accuracy \citep[][using the caret package by]{caret}. The best performing model based on balanced accuracy and number of false negatives (incorrectly predicted `No Fires') is reported in Table~\ref{table:logf}.

@@

%ROC for logits
\begin{figure}
  \centering
	\includegraphics[width=.8\textwidth]{figures/all_roc.pdf}
  \caption{ROC curve for logistic models}
  \label{fig:ffdi}
\end{figure}

@@

As classification trees can be sensitive to the sample used, we constructed several trees on different balanced samples of the data. As each of the trees were structured very similarly we only report two in Figure~\ref{fig:tree2}.

% Trees with all variables
\begin{figure}
  \centering
	\includegraphics[page=1, width=\textwidth]{figures/trees2.pdf}\par\vspace*{1.5cm}\par
	\includegraphics[page=2, width=\textwidth]{figures/trees2.pdf}
  \caption{Trees with all variables}
  \label{fig:tree2}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=.8\textwidth]{figures/trees_roc.pdf}
  \caption{ROC curve for Tree Models}
  \label{fig:troc}
\end{figure}

\subsection{Random Forest}

We produce the random forest using the caret and randomForest packages \citep{caret, rf} in R. The random forest is built using the balanced training set and then assessed on its prediction on the test set.

@@

\begin{figure}
  \centering
  	\includegraphics[width=.7\textwidth]{figures/varimp.pdf}
  \caption{Variable Importance for random forest}
  \label{fig:vimp}
\end{figure}

@@

\begin{figure}
  \centering
	\includegraphics[width=.8\textwidth]{figures/roc_rf.pdf}
  \caption{Variable Importance for random forest}
  \label{fig:rocrf}
\end{figure}

@@

This thesis takes a very simplistic view of the dataset, rather than utilising the spatial and time series component, we treat each spatial and time point as independent. Utilising the lagged variables will no doubt improve the model for example we'd expect after a week of hot windy days the probability of ignition would be higher compared to a week of cold days or rain.

@@

From the different models in this paper, we conclude that the FFDI alone of a poor predictor of bushfire ignition.
